A full approximation scheme multilevel method \\ for nonlinear variational inequalities

### 2023-07-14 paper/fascd.tex
clean Section 3; having floats for the basic CD iterations is not necessary

Suppose that there are $m<\infty$ closed subspaces $\cV_i \subset \cV$, so that the sum
\begin{equation}
\cV = \sum_{i=0}^{m-1} \cV_i \label{eq:subspacedecomp}
\end{equation}
holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.  We say \eqref{eq:subspacedecomp} is a \emph{space decomposition} of $\cV$ \cite{Xu1992}.  For a closed, convex subset $\cK \subset \cV$, suppose further that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
\begin{equation}
\cK = \sum_{i=0}^{m-1} \cK_i. \label{eq:constraintdecomp}
\end{equation}
The sum in \eqref{eq:constraintdecomp} must hold in two senses \cite{TaiTseng2002}: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.  Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique, and also notice that sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.  Finally, suppose there are bounded \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.  Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:femultilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
\begin{equation}
v = \sum_{i=0}^{m-1} \Pi_i v.  \label{eq:constraintrestrictionsum}
\end{equation}
The operators $\Pi_i$ are generally nonlinear.  Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}. Tai \cite{Tai2003} defined a \emph{constraint decomposition} (CD) of $\cK$ to be a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}.  Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; note $\cK_i \not\subset \cK$ in this, and in most nontrivial cases as well.

@@

In Sections \ref{sec:femultilevel}--\ref{sec:vcycle} we will introduce multilevel CD discretizations and algorithms for finite-dimensional VI problems.  However, one can observe here that the CD concept applies at the level of the continuum problem, as illustrated in the following two examples, the first of which is an overlapping domain decomposition of a Sobolev space.  Note that $\cK_i \not\subset \cK$ in these cases; informally speaking, the important inclusion is $\cK_i \subset \cV_i$.

\begin{example}  \label{ex:domaindecomposition}  For a bounded domain $\Omega \subset \RR^d$, let $\cV = W_0^{k,p}(\Omega)$ for $k\ge 0$ and $p\ge 1$.  Let $\psi$ be a measurable, real-valued obstacle, defined on $\Omega$, which satisfies $\psi|_{\partial \Omega} \le 0$, and let $\cK = \{v \ge \psi\} \subset \cV$.  Suppose further that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, satisfying $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the open support of $\phi_i$.  Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in \cV_i: v \ge \phi_i \psi\}$, and $\Pi_i(v) = \phi_i v$.  Then \eqref{eq:subspacedecomp}, \eqref{eq:constraintdecomp}, and \eqref{eq:constraintrestrictionsum} all hold.
\end{example}

Our second example is a disjoint frequency decomposition of a Hilbert space.  A multilevel CD over a finite-dimensional FE discretization, as proposed in Sections \ref{sec:femultilevel} and \ref{sec:cdmultilevel}, approximates such a decomposition.

\begin{example}  \label{ex:frequencydecomposition}  For simplicity suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = W_{\text{per}}^{2,k}(\Omega)$ be the periodic functions with square-integrable $k$th derivatives for $k \ge 0$.  Suppose $\psi \in W_{\text{per}}^{2,k}(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.  Without using any detailed notation for Fourier representation, but noting that the frequencies are discrete because $\Omega$ is compact, suppose $\{\cV_i\}$ are $m<\infty$ subspaces of $\cV$ defined by an (nonoverlapping) partition of $W_{\text{per}}^{2,k}(\Omega)$ by frequency, thus satisfying \eqref{eq:subspacedecomp} as an orthogonal decomposition.  Suppose $P_i:\cV \to \cV_i$ are the corresponding orthogonal projections, satisfying $I = \sum_i P_i$.  Let $\cK_i = \{v \ge P_i \psi\} \subset \cV_i$ and $\Pi_i = P_i$.  Then \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} also hold.
\end{example}

Associated to a constraint decomposition \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} are certain iterative methods \cite{Tai2003,Xu1992}.  Let $0\le \alpha \le 1$ be the \emph{damping parameter}.\footnote{FAScd Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, and indeed it seems to be unnecessary.  However, the convergence of the more general CD iterations appears to depend upon it.}  Algorithms \ref{alg:basiccd-add} and \ref{alg:basiccd-mult} below, discussed next, solve smaller VI problems over each subset $\cK_i$.  Each Algorithm replaces $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}.

The additive/parallel \pr{cd-add} Algorithm \ref{alg:basiccd-add} generalizes the classical Jacobi iteration.
\begin{pseudofloat}[H]
\begin{pseudo*}
\pr{cd-add}(u)\text{:} \\+
    for all $i \in \{0,\dots,m-1\}$: \\+
        \rm{find} $\hat w_i\in \cK_i$ \rm{such that} \\+
            $\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \cK_i$ \\--
    $\hat w = \sum_i \hat w_i\in\cK$ \\
    return $w=(1-\alpha) u + \alpha \hat w$
\end{pseudo*}
\caption{One additive CD iteration for VI problem \eqref{eq:vi}.}
\label{alg:basiccd-add}
\end{pseudofloat}
Here ``\textbf{for all}'' can be computed in any order.  In its VI,
\begin{equation}
\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \mathcal{K}_i, \label{eq:cdaddvi}
\end{equation}
the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.
%\pef{Could remove this explanation, but it's nice to keep it in if we can fit it}
% Proposing to keep for now.

The multiplicative/successive \pr{cd-mult} Algorithm \ref{alg:basiccd-mult} generalizes the Gauss--Seidel iteration.  \begin{pseudofloat}[H]
\begin{pseudo*}
\pr{cd-mult}(u)\text{:} \\+
    for $i = 0,\dots,m-1$: \\+
        \rm{find} $\hat w_i\in \cK_i$ \rm{such that} \\+
            $\displaystyle \ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \cK_i$ \\-
            $w_i = (1-\alpha) \Pi_i u + \alpha \hat w_i\in\cK_i$ \\-
    $\hat w = \sum_i \hat w_i\in\cK$ \\
    return $w=(1-\alpha) u + \alpha \hat w$
\end{pseudo*}
\caption{One multiplicative CD iteration for VI problem \eqref{eq:vi}.}
\label{alg:basiccd-mult}
\end{pseudofloat}
Each subset solution corrects the global iterate immediately, so here the ordering used in the ``\textbf{for}'' loop is important.  In the VI,
\begin{equation}
\ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \qquad \forall v_i \in \mathcal{K}_i, \label{eq:cdmultvi}
\end{equation}
the sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those portions of $u$ which have not yet improved''.

@@

We call \pr{cd-add} and \pr{cd-mult} the \emph{basic CD iterations}.  VI problems \eqref{eq:cdaddvi} and \eqref{eq:cdmultvi} reduce to constrained line searches when the $\cV_i$ are one-dimensional.  As shown in the next Lemma, when VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then, subject to bounds on the damping $\alpha$, these iterations generate monotonically-decreasing objective values.
\begin{lemma} \cite{Tai2003}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0 \le \alpha \le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
\begin{equation}
F(w) - \ip{\ell}{w} \le F(u) - \ip{\ell}{u}.  \label{eq:objectivemonotone}
\end{equation}
If $0 \le \alpha \le 1/m$ then \pr{cd-add} likewise satisfies \eqref{eq:objectivemonotone}.
\end{lemma}

@@

The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and even when $f$ is nonlinear, non-local, or defined only on $\cK$.  However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.  References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$. Substantial implementation simplifications are available in such cases.

\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear and is local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.  Considering \pr{cd-add} for concreteness, by linearity \eqref{eq:cdaddvi} can be rewritten
\begin{equation}
\ip{f(e_i)}{v_i-\hat w_i} \ge \ip{\tilde\ell}{v_i-\hat w_i}, \label{eq:linearlocalvi}
\end{equation}
for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$ incorporates the initial residual $f(u)$.  Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible---generally $e_i \notin \cK$---but \eqref{eq:linearlocalvi} is meaningful for this $f$.
\pef{Can you expand a little bit here? I confess I don't follow.}
\end{example}

When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.  Section \ref{sec:vcycle} constructs a multiplicative multilevel CD iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}.

From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.  Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.  The \emph{lower obstacle} $\underline{\gamma}$ and the \emph{upper obstacle} $\overline{\gamma}$ are given functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$.  We suppose $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$ and $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$ are measurable, $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$, and $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$, and that $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.  Dirichlet data\footnote{Dirichlet boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} $g_D:\partial_D \Omega \to \RR$ then defines
\begin{equation}
\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega), \label{eq:originalconstraintset}
\end{equation}
the closed and convex \emph{constraint set}.  Note that we are treating inequality constraints and boundary conditions in a unified manner.\footnote{Any Neumann conditions over $\partial_N \Omega$ are imposed weakly in the definition of $f$, i.e.~by an integration over $\partial_N\Omega$.}

We recall the abstract VI problem \eqref{eq:vi} with constraint sets \eqref{eq:originalconstraintset}.  Given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, we will assume that the following is well-posed for $u^\star\in \cK$:
\begin{equation}
\ip{f(u^\star)}{v-u^\star} \ge \ip{\ell}{v-u^\star} \qquad \text{for all } v\in \cK. \label{eq:boxdirichletvi}
\end{equation}
Note that $v-u^\star \in W_0^{1,p}(\Omega)$.

@@

We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:boxdirichletvi} using finite elements (FE).  Let $\Omega \subset \RR^d$ be a (open) bounded polygon. We will construct a multilevel FE approximation to \eqref{eq:boxdirichletvi} over $\Omega$ using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a nested mesh hierarchy; the goal is an accurate solution on the finest-level mesh.  The hierarchy allows the creation of a multilevel CD of $\cK$, using the ideas in the next two Sections, which permits solvers based on admissibility-preserving coarse corrections.

@@

Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.  Here there is, in fact, another ``basic'' iteration, an incomplete variant of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}) that does not need the $\Pi_i$ operators.  It is a variant of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}).

\begin{pseudofloat}[ht]
\begin{pseudo*}
\pr{icd-tele}(w^J)\text{:} \\+
    for $j = 0,\dots,J$: \\+
        \rm{find} $z^j\in \mathcal{U}^j$ \rm{so that for all} $v \in \mathcal{U}^j$, \\+
            $\displaystyle \ip{f\Big(w^J + \alpha \sum_{i<j} z^i + \alpha z^j\Big)}{v-z^j} \ge \ip{\ell}{v-z^j}$ \\--
    return $\tilde w^J=w^J + \alpha \sum_{i\le J} z^i$
\end{pseudo*}
\caption{One telescoping incomplete CD iteration for VI problem \eqref{eq:fe:vi}, assuming $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.}
\label{alg:basiccd-tele}
\end{pseudofloat}

@@

Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ by incorporating information from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.}  The corrections are constructed, inductively, so that iterates remain admissible at every level.  Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  Solving the coarsest-level VI results in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way (Algorithm \ref{alg:basiccd-tele}), up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.  Then the next finest-level, admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.

>>>>

Suppose that there are $m<\infty$ closed subspaces $\cV_i \subset \cV$, so that the sum
\begin{equation}
\cV = \sum_{i=0}^{m-1} \cV_i \label{eq:subspacedecomp}
\end{equation}
holds in the sense that if $w \in \cV$ then there exist $w_i \in \cV_i$ so that $w = \sum_i w_i$.  We say \eqref{eq:subspacedecomp} is a \emph{space decomposition} of $\cV$ \cite{Xu1992}.  For a closed, convex subset $\cK \subset \cV$, suppose further that $\cK_i \subset \cV_i$ are nonempty, closed, and convex subsets such that
\begin{equation}
\cK = \sum_{i=0}^{m-1} \cK_i. \label{eq:constraintdecomp}
\end{equation}
The sum in \eqref{eq:constraintdecomp} must hold in two senses \cite{TaiTseng2002}: \emph{(i)}~if $w \in \cK$ then there exist $w_i \in \cK_i$ so that $w = \sum_i w_i$, and \emph{(ii)}~if $z_i \in \cK_i$ for each $i$ then $\sum_i z_i \in \cK$.  Note that neither decomposition \eqref{eq:subspacedecomp} or \eqref{eq:constraintdecomp} is required to be unique, and also notice that sense \emph{(ii)} is automatic for \eqref{eq:subspacedecomp} because the $\cV_i$ are subspaces.  Finally, suppose there are bounded, generally nonlinear, \emph{decomposition operators}\footnote{Denoted ``$R_i$'' in \cite{Tai2003}.  Here we avoid a notational conflict with multilevel transfer operators (Section \ref{sec:femultilevel}).} $\Pi_i : \cK \to \cK_i$ such that if $v \in \cK$ then
\begin{equation}
v = \sum_{i=0}^{m-1} \Pi_i v.  \label{eq:constraintrestrictionsum}
\end{equation}
Clearly \eqref{eq:constraintrestrictionsum} implies sense \emph{(i)} for \eqref{eq:constraintdecomp}.  A \emph{constraint decomposition} (CD) of $\cK$ is a choice of $\cV_i,\cK_i,\Pi_i$ satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} \cite{Tai2003}.  Figure \ref{fig:cartoon} suggests how a low-dimensional example might look; note $\cK_i \not\subset \cK$ in this and most other cases.

@@

In Sections \ref{sec:femultilevel}--\ref{sec:vcycle} we will introduce multilevel FE CD discretizations and algorithms for finite-dimensional VI problems.  However, the CD concept applies at the level of the continuum problem, as illustrated in the following two examples.  The first example is an overlapping domain decomposition of a Sobolev space.  Again note that $\cK_i \not\subset \cK$; informally speaking, the important inclusion is $\cK_i \subset \cV_i$.

\begin{example}  \label{ex:domaindecomposition}  For a bounded domain $\Omega \subset \RR^d$, let $\cV = W_0^{k,p}(\Omega)$ for $k\ge 0$ and $p\ge 1$.  Let $\psi$ be a measurable, real-valued obstacle, defined on $\Omega$, which satisfies $\psi|_{\partial \Omega} \le 0$, and let $\cK = \{v \ge \psi\} \subset \cV$.  Suppose that $\{\phi_i\}_{i=0}^{m-1}$ is a smooth partition of unity on $\Omega$, satisfying $0 \le \phi_i\le 1$ and $\sum_i \phi_i = 1$, and let $\Omega_i$ be the open support of $\phi_i$.  Let $\cV_i = \{w \in \cV:w|_{\Omega \setminus \Omega_i} =0 \}$, $\cK_i = \{v \in \cV_i: v \ge \phi_i \psi\}$, and $\Pi_i(v) = \phi_i v$.  Then \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum} hold, giving a CD.
\end{example}

Our second example is a disjoint frequency decomposition of a Hilbert space.  A multilevel CD over a finite-dimensional FE discretization (Sections \ref{sec:femultilevel}--\ref{sec:cdmultilevel}) should approximate this CD.

\begin{example}  \label{ex:frequencydecomposition}  Suppose $\Omega = (0,a)^d \subset \RR^d$ is a cube, and let $\cV = W_{\text{per}}^{2,k}(\Omega)$ be the periodic functions with square-integrable $k$th derivatives, for $k \ge 0$.  Suppose $\psi \in W_{\text{per}}^{2,k}(\Omega)$ and let $\cK = \{v \ge \psi\} \subset \cV$.  Without using any detailed notation for Fourier representation, but noting that the frequencies are discrete because $\Omega$ is compact, suppose $\{\cV_i\}$ are $m<\infty$ subspaces of $\cV$ defined by an (nonoverlapping) partition of $W_{\text{per}}^{2,k}(\Omega)$ by frequency, thus satisfying \eqref{eq:subspacedecomp} as an orthogonal decomposition.  Suppose $P_i:\cV \to \cV_i$ are the corresponding orthogonal projections, satisfying $I = \sum_i P_i$.  Let $\cK_i = \{v \ge P_i \psi\} \subset \cV_i$ and $\Pi_i = P_i$.  Then \eqref{eq:constraintdecomp} and \eqref{eq:constraintrestrictionsum} also hold.
\end{example}

Associated to any CD are certain iterative methods \cite{Tai2003,Xu1992}.  Let $0\le \alpha \le 1$ be the \emph{damping parameter}.\footnote{FAScd Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} are presented without damping, and indeed it seems to be unnecessary.  However, the convergence of the general CD iterations appears to depend upon it.}  The following algorithms, which replace $u \in \cK$ with a new iterate $w\in\cK$ which should be closer to $u^\star \in \cK$ solving \eqref{eq:vi}, solve smaller VI problems over each subset $\cK_i$.

The additive/parallel \pr{cd-add} algorithm generalizes the classical Jacobi iteration.  In its VI problem the expression ``$u-\Pi_iu+\hat w_i$'' removes the part of $u$ which lies in $\mathcal{K}_i$, and then adds-back the subset solution $\hat w_i \in \mathcal{K}_i$.

\begin{pseudo*}
\pr{cd-add}(u)\text{:} \\+
    for all $i \in \{0,\dots,m-1\}$: \\+
        \rm{find} $\hat w_i\in \cK_i$ \rm{such that} \\+
            $\ip{f(u - \Pi_i u + \hat w_i)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \cK_i$ \\--
    $\hat w = \sum_i \hat w_i\in\cK$ \\
    return $w=(1-\alpha) u + \alpha \hat w$
\end{pseudo*}

The multiplicative/successive \pr{cd-mult} algorithm generalizes the Gauss--Seidel iteration.  Each subset solution corrects the global iterate immediately, so here the ordering of sets $\{\cK_i\}$ is important.  In its VI problem the sum ``$\sum_{j>i} \Pi_j u$'' should be read as ``those portions of $u$ which have not yet improved''.

\begin{pseudo*}
\pr{cd-mult}(u)\text{:} \\+
    for $i = 0,\dots,m-1$: \\+
        \rm{find} $\hat w_i\in \cK_i$ \rm{such that} \\+
            $\displaystyle \ip{f\Big(\sum_{j<i} w_j + \hat w_i + \sum_{j>i} \Pi_j u\Big)}{v_i-\hat w_i} \ge \ip{\ell}{v_i-\hat w_i} \textnormal{ for all } v_i \in \cK_i$ \\-
            $w_i = (1-\alpha) \Pi_i u + \alpha \hat w_i\in\cK_i$ \\-
    $\hat w = \sum_i \hat w_i\in\cK$ \\
    return $w=(1-\alpha) u + \alpha \hat w$
\end{pseudo*}

@@

We call \pr{cd-add} and \pr{cd-mult} the \emph{basic CD iterations}.  The VI problems solved in these iterations reduce to constrained line searches when the $\cV_i$ are one-dimensional.  When VI problem \eqref{eq:vi} corresponds to convex optimization \eqref{eq:minimization} then, subject to bounds on the damping $\alpha$, these iterations generate monotonically-decreasing objective values.
\begin{lemma} \cite{Tai2003}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0 \le \alpha \le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
\begin{equation}
F(w) - \ip{\ell}{w} \le F(u) - \ip{\ell}{u}.  \label{eq:objectivemonotone}
\end{equation}
If $0 \le \alpha \le 1/m$ then \pr{cd-add} likewise satisfies \eqref{eq:objectivemonotone}.
\end{lemma}

@@

The basic CD iterations are meaningful whether or not we are in the optimization case $f=F'$, and when $f$ is nonlinear, non-local, or defined only on $\cK$.  However, practical and efficient FE implementation for such general operators $f$ seems not to have been addressed in the literature.  References \cite{GraeserKornhuber2009,Tai2003} only apply the CD method to the classical obstacle problem, wherein $f=F'$ is linear, local, and defined on all of $\mathcal{V}$. Substantial implementation simplifications are available in such cases.

\begin{example}  \label{ex:fnice} Suppose $f:\cV \to \cV'$ is linear, and suppose it is local in the sense that for any $z\in\mathcal{V}$ the value $\ip{f(\phi)}{z}$ can be computed by an integral over the support of $\phi$.  Considering \pr{cd-add} for concreteness, by linearity the VI can be rewritten
\begin{equation}
\ip{f(e_i)}{v_i-\hat w_i} \ge \ip{\tilde\ell}{v_i-\hat w_i}, \label{eq:linearlocalvi}
\end{equation}
for all $v_i \in \mathcal{K}_i$, where $\tilde\ell = \ell - f(u)$ incorporates the initial residual $f(u)$.  Note $e_i = \hat w_i - \Pi_i u \in \cV_i$ is not generally admissible---generally $e_i \notin \cK$---but ``$f(e_i)$'' in \eqref{eq:linearlocalvi} is meaningful because $f$ is defined on all of $\cV$.
\end{example}

%\pef{Can you expand a little bit here? I confess I don't follow.}
% Tried

When $f$ has the locality of Example \ref{ex:fnice}, an FE approximation of VI \eqref{eq:vi} can be solved in any incremental and efficient manner following the techniques in \cite{GraeserKornhuber2009,Tai2003}, namely by computations over basis function supports.\footnote{Compare equations (2.31) and (2.32) in \cite{Farrelletal2021}, which describe the same sense of locality.}  However, for general problems lacking the niceties of Example \ref{ex:fnice}, certain additional ideas are needed for practical implementation of CD-type iterations.  Section \ref{sec:vcycle} constructs such an iteration for a general operator $f:\cK\to\cV'$ by applying the full approximation scheme (FAS) idea of Brandt \cite{Brandt1977}.

From now on we restrict to the case of \emph{box constraints} \cite{BensonMunson2006,FerrisPang1997}, equivalently \emph{bi-obstacle problems}.  Suppose $\Omega \subset \RR^d$ is open and let $\mathcal{V}=W^{1,p}(\Omega)$.  The \emph{lower obstacle} $\underline{\gamma}$ and \emph{upper obstacle} $\overline{\gamma}$ are given measurable functions, defined on the closure $\overline{\Omega}$, with values in the extended real line $\tilde\RR = [-\infty,+\infty]$, satisfying the following properties: $\underline{\gamma} : \overline{\Omega} \to [-\infty,+\infty)$, $\overline{\gamma} : \overline{\Omega} \to (-\infty,+\infty]$, and $\underline{\gamma} \le \overline{\gamma}$ on $\overline{\Omega}$.  Also suppose $\partial\Omega$ is split into disjoint, measurable sets, $\partial\Omega = \partial_D \Omega \cup \partial_N \Omega$, with $\partial_D \Omega$ of positive Hausdorff measure, and $\partial_N \Omega$ possibly empty.  Dirichlet data\footnote{Dirichlet boundary values are in the trace sense \cite{Evans2010}, and $g_D$ is assumed to be as regular as needed.} is given by $g_D:\partial_D \Omega \to \RR$, and we assume $\underline{\gamma} \le g_D \le \overline{\gamma}$ on $\partial_D \Omega$.  Now define
\begin{equation}
\cK = \left\{w\,:\,\underline{\gamma} \le w \le \overline{\gamma} \, \text{ and }\, w\big|_{\partial_D \Omega} = g_D\right\} \subset \cV =W^{1,p}(\Omega), \label{eq:originalconstraintset}
\end{equation}
the closed and convex \emph{constraint set}.  Note that we are treating inequality constraints and Dirichlet boundary conditions in a unified manner.  Any Neumann conditions over $\partial_N \Omega$ are imposed weakly in the definition of the operator $f$, i.e.~by an integration over $\partial_N\Omega$.

In the next Sections we numerically solve continuum VI problem \eqref{eq:vi} over a constraint set of the form \eqref{eq:originalconstraintset}.  In fact, given continuous $f:\cK \to \cV'$, and $\ell \in \cV'$, from now on we assume that the following is well-posed for $u^\star\in \cK$:
\begin{equation}
\ip{f(u^\star)}{v-u^\star} \ge \ip{\ell}{v-u^\star} \qquad \text{for all } v\in \cK. \label{eq:boxdirichletvi}
\end{equation}
Note that $v-u^\star \in W_0^{1,p}(\Omega)$.

@@

We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:boxdirichletvi} using finite elements (FE).  Let $\Omega \subset \RR^d$ be a (open) bounded polygon.  A multilevel FE approximation to \eqref{eq:boxdirichletvi} is constructed, using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a mesh hierarchy.  The hierarchy allows the creation of a multilevel CD of $\cK$ using the ideas in the next two Sections.

@@

Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.  Here there is, in fact, another ``basic'' iteration, an incomplete variant of \pr{cd-mult} (Section \ref{sec:cd}) that does not need the $\Pi_i$ operators.  One application of \pr{icd-tele} below assumes $w^J\in\mathcal{K}^J$, property \eqref{eq:fe:finestlevelequivalent}, and incomplete CD \eqref{eq:fe:upwardsuminclusion}.

\begin{pseudo*}
\pr{icd-tele}(w^J)\text{:} \\+
    for $j = 0,\dots,J$: \\+
        \rm{find} $z^j\in \mathcal{U}^j$ \rm{so that for all} $v \in \mathcal{U}^j$, \\+
            $\displaystyle \ip{f\Big(w^J + \alpha \sum_{i<j} z^i + \alpha z^j\Big)}{v-z^j} \ge \ip{\ell}{v-z^j}$ \\--
    return $\tilde w^J=w^J + \alpha \sum_{i\le J} z^i$
\end{pseudo*}

@@

Each V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ by incorporating information from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections all live in the fine-level subspace $\mathcal{V}^J$.\footnote{Nonetheless we use prolongation $P$ to indicate where the computer representation changes.}  The corrections are constructed, inductively, so that iterates remain admissible at every level.  Suppose $y^i \in \mathcal{D}^i$, for $i=J,\dots,j+1$, are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).  By \eqref{eq:fe:downwardsuminclusion} the correction at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  Solving the coarsest-level VI results in the full downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$, which then yields $z^1 \in \mathcal{U}^1$ and an admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$, using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the V-cycle correction.  Then the next finest-level, admissible iterate is $\tilde{w}^J = w^J + z^J \in \mathcal{K}^J$.
